### 2022-03-30 paper/Chapters/Background.tex
(almost) finish draft of hardware section
Still need to investigate the USB issues and maybe add more detail to the end of the chapter, but other than that I have a couple of questions to resolve then I'm happyish with the chapter

All computer products are a balance, it simply isn't possible to have the perfect combination of portable, powerful, serviceable, and cheap.
While desktops focus on power and serviceability, laptops often focus portability at the expense of other traits\todo{Theres a word at the tip of my tounge thats a lot better here dangit}.
As mobility becomes ever more important, gaming laptops have become more popular to bring power to the portable form factor.
This understandably comes with a cost, with the literal cost of a gaming laptop being the most obvious compromise.
The componentization of such a laptop also often takes a hit with parts being directly soldered to the motherboard, reducing a user's options for repairs or upgrades.
It is more expensive, and often less user-serviceable, to buy a gaming laptop that has the same performance of a desktop computer -- but given that it is an all-in-one device, it's portability gives it a niche that a desktop cannot hope to rival.
The compromise presented by a gaming laptop helps close the gap between power and portability, but it still falls short of being an all around good solution.

>>>>

All computer products are a balance, it simply isn't possible to have the perfect combination of portable, powerful, serviceable, and cheap.
While desktops focus on power and serviceability, laptops often focus portability at the expense of other attributes.
As mobility becomes ever more important, gaming laptops have become more popular to bring power to the portable form factor.
This understandably comes with a cost, with the literal cost of a gaming laptop being the most obvious compromise.
The componentization of such a laptop also often takes a hit with parts being directly soldered to the motherboard, reducing a user's options for repairs or upgrades.
It is more expensive, and often less user-serviceable, to buy a gaming laptop that has the same performance of a desktop computer -- but given that it is an all-in-one device, it's portability gives it a niche that a desktop cannot hope to rival.
The compromise presented by a gaming laptop helps close the gap between power and portability, but it still falls short of being an all around good solution.

### 2022-03-30 paper/Chapters/Hardware.tex
(almost) finish draft of hardware section
Still need to investigate the USB issues and maybe add more detail to the end of the chapter, but other than that I have a couple of questions to resolve then I'm happyish with the chapter

Immediately, the device must be capable of decoding and rendering video that is streamed over the internet.
Such video processing is a relatively expensive process that cannot easily be done with something such as a micro controller or even a single core CPU seen on Single Board Computers (SBC) such as the Raspberry Pi Zero due to the lack of working memory and processing power \cite{picockpit_2021}.
\todoadvisor{Add more detail?}
Though at the same time, it is important to not simply build a powerful system akin to a laptop that will consume more power.
It makes no sense to built a powerful mini-computer when all it will ever be used to is to connect to another computer, and the added power draw would require a larger battery in order to stay operational for decent periods of time.
Instead it should be cut down and distilled into something that can do it's job as efficiently as possible.

@@

This leads to the desire for a portable design that can be used both in a conducive environnement such as a desk or while mobile.
The hardware should have the ability to be powered by a chord plugging into the device, or by an internal battery.
This alone presents a number of difficult concerns, but those will be addressed later in this chapter\todo{cite}.
Such a requirement leads to a clear benefit of integrating crucial computers such as a battery and a screen into the hardware design, while leaving as many ports open such as ethernet and USB for the user to take advantage of.

@@

@@

\todo{Add paragraph on DSI vs HDMI}

\todo{Add paragraph RTC}

@@

\todosection

>>>>

Immediately, the device must be capable of decoding and rendering video that is streamed over the internet.
Such video processing is a relatively expensive process that cannot easily be done with something such as a micro controller or even a single core CPU seen on Single Board Computers (SBC) such as the Raspberry Pi Zero due to the lack of working memory and processing power \cite{picockpit_2021}.
Though at the same time, it is important to not simply build a powerful system akin to a laptop that will consume more power.
It makes no sense to built a powerful mini-computer when all it will ever be used to is to connect to another computer, and the added power draw would require a larger battery in order to stay operational for decent periods of time.
Instead it should be cut down and distilled into something that can do it's job as efficiently as possible.

@@

This leads to the desire for a portable design that can be used both in a conducive environnement such as a desk or while mobile.
The hardware should have the ability to be powered by a chord plugging into the device, or by an internal battery.
This alone presents a number of difficult concerns which will be addressed in Section \ref{sec:Prototyping}.
Such a requirement leads to a clear benefit of integrating crucial computers such as a battery and a screen into the hardware design, while leaving as many ports open such as ethernet and USB for the user to take advantage of.

@@

Thought must also be given to the peripherals that will be included in the design.
Most importantly, if a screen is going to be included in the design, what kind of screen will be used?
For projects of this scale, the usual candidates for a relatively performant screen are HDMI screens or DSI screens.
While HDMI screens are much more common and recognizable, DSI screens are a new development that promises high performance in low cost and low power applications such as embedded systems.
However, when looking into existing DSI screens, there are a couple of issues that prevent it from being the best choice for this project.
Even though DSI screens are more power efficient and have built in touch support, it is a closed format which greatly restricts manufactures from producing their own screens \cite{dsi_vs_hdmi}.
Due to this, there is only one screen available for purchase outside of the official implementation, and both have a lower resolution that what is possible with HDMI screens of the same size.
So even though HDMI screens are more expensive at this size, they are a more reliable choice compared to the still-developing DSI standard.

Another potential feature to consider is whether to integrate a Real Time Clock (RTC) into the design.
An RTC is a component commonly found within laptop and desktop computers that is used to keep track of time even though the computer is powered off.
This is done with a small button battery powering an RTC chip integrated into the motherboard to keep the system's clock in sync with the real world.
While this is almost never necessary in embedded systems, it is worth considering for this project because of it's use as a mobile computer.
However, when considering the primary use case of connecting to another computer over the internet, and the potential of needing a linux kernel patch in order to add support for the RTC \cite{rtc_kernelpatch}, it would be much simpler to tell the operating system to simply resync it's clock with an internet clock every time it connects to a network.

@@

@@

\begin{figure}[b]
  \centering
  \includegraphics[width=0.5\textwidth]{Figures/kicad/close-ups/stencil}
  \caption[PCB Stencil]{The PCB's stencil. Areas colored black are where solder paste is applied to solder SMD parts to the board}
  \label{fig:PCBStencil}
\end{figure}

\todo[inline]{Insert pictures of PCB. Thinking at least one showing the flaws of the first board, another of the completed board}

Once the board's design has been completed and checked against all electrical, functional, and manufacturing constraints, the Printed Circuit Board can be produced.
The first step in fabrication is to print the actual PCB.
While ASU offers a basic printing service for use in classes and academic projects, the requirements of this board, specifically the tiny traces and pads required for the CM4's mezzanine connectors (pictured in Figure \ref{fig:rpi_cm4_mezzanine}), are outside of the printer's capabilities.
Instead, by recommendation of a professor, the printing was done through JLCPC (\href{https://www.jlcpc.com/}{JLCPCB.com})\todoquestion{Would this be the right format for linking JLCPCB? It's not really a source or a reference that would be linked in the bibliography}, a company that provides quick turnaround and low costs for printing PCBs.
A set of five boards was ordered, allowing for multiple boards to be assembled in case of error, as well as a stencil.
A PCB stencil is a thin sheet of metal with holes cut in it that line up with the pads on the PCB board.
This enables the easy application of solder paste onto all of the pads on the board as once, including the difficulty small pads of the CM4's mezzanine connectors.
While the PCB was being printed, all of the parts found in the Bill of Materials (found in Appendix \ref{AppendixB}) were ordered from sites such as Digi-Key and Mouser.\todoquestion{Should I mention small issues such as the time when I ordered the PCB and before I could order the parts one of them went out of stock? I worry that that would break the flow of the paper, and I already mentioned stock issues in a previous section. I feel like it's a detail that makes an interesting story, but since this isn't a story paper it's not needed here.}
In order to account for the production of multiple prototype boards and the possibility of broken parts, enough parts were ordered to produce five complete boards.

Once the PCB and it's component parts arrived, it was time for assembly.
The first board was assembled by hand using the PCB stencil and a handheld heat gun.
The stencil, shown as an image in Figure \ref{fig:PCBStencil}, is used by securing it the PCB board so that the holes line up with the pads, and spreading solder paste across it allowing the paste to cover the exposed pads through the holes.
The stencil was then removed, and each of the surface mounted components were then placed onto the board using a set of tweezers.
For tricky parts such as the CM4's mezzanine connectors and other small parts with the pads below the component itself, a microscope was used to assist in lining up the pads with the components.
Once the parts were placed, a heat gun was used to melt the solder paste and fuse the components to the board.
Following the SMDs, the through-hole components, such as the ports lining the side of the board, were then placed and soldered using a hand-held \todoquestion{soldering iron or solder iron?}soldering iron.

The first board ran into two issues that prevented it from working.
Firstly, due to improper use of the heat gun, most of the LEDs had popped, resulting in all hardware-level diagnostic indicators being inoperable.
Secondly, the CM4 connectors were not properly connected to the board, and in an attempt to repair it, the connector components were irreparably damaged and the board itself was damaged slightly.\todo{Definitely want a picture here}
Instead of attempting to assemble another board by hand, a local third-party company was contacted to assemble the board professionally.
After delivering the board, stencil, and parts, the second board was assembled.

\begin{figure}[b]
  \centering
  \begin{subfigure}{.5\textwidth}
    \centering
    \includegraphics[width=.8\linewidth]{Figures/kicad/QFN-20-1EP_4x4mm}
    \caption{\emph{QFN-20-1EP\_4x4mm}}
    \label{fig:FootprintQFN204x4_Original}
  \end{subfigure}%
  \begin{subfigure}{.5\textwidth}
    \centering
    \includegraphics[width=.8\linewidth]{Figures/kicad/QFN-20-1EP_4x4mm_RevA}
    \caption{\emph{QFN-20-1EP\_4x4mm\_RevA}}
    \label{fig:FootprintQFN204x4_RevA}
  \end{subfigure}
  \caption[Two versions of the QFN-20-1EP\_4x4mm package]{The two footprints for components with the \emph{QFN-20-1EP\_4x4mm} package. The original footprint is shown in Figure \ref{fig:FootprintQFN204x4_Original}, while the revised footprint is shown in Figure \ref{fig:FootprintQFN204x4_RevA}}
  \label{fig:FootprintQFN204x4}
\end{figure}

The second board was assembled mostly successfully, and allowed the first stage of testing to begin.
The only issue apart from some ports being too close together with real-world cables was one component having the wrong footprint on the board.
Unfortunately, the component \emph{U2 (MCP73871-2CCI\_ML)}, is described as using the \emph{QFN-20-1EP\_4x4mm} footprint in it's datasheet\todoquestion{Does this need to be cited? Can you cite a datasheet like this?}, but actually uses a revised version of the footprint, shown in Figure \ref{fig:FootprintQFN204x4}.
This discrepancy meant that that component could not be soldered, and it's associated circuit would not function properly.
In this case, the component was a part of the battery charging circuit, causing the entire primary power supply to be unusable.
Thankfully, as described in Section \ref{sec:Prototyping}, the secondary method of powering the board could be used to test the remaining features of the board.
After making the necessary fixes, a second round of five PCB boards and a stencil was ordered from JLCPCB.
Because there was still enough components to assemble three more boards, and no components needed to be changed, no more components were ordered at this stage.
Once the boards were delivered, the revised board was again sent to be professionally assembled.

Once the third board was assembled, the board could be tested in full.
While each part of the board was at least partially functional, there were still issues preventing it from working completely.
In particular, USB devices plugged into any of the ports on the board were able to receive power from the board but were unable to transfer data in either direction.
\todo[inline]{I am currently going through and diagnosing this issue}
